---
title: "0.1_monthly_chla_download"
author: "mz"
format: html
editor: source
---

## Libraries

load the packages needed for downloading the data

```{r}


library(RCMEMS)
library(raster)
```

load credentials


```{r}
# read cred file ingoring row names


cred <- read.csv(here::here('data/credentials.csv'), sep = ";", row.names = F)

user <- cred$user
pss <- cred$pss

```

```{r}

```
(raw) Script to download files

```{r}
scriptGEN <- 'python -m motuclient --motu https://my.cmems-du.eu/motu-web/Motu --service-id OCEANCOLOUR_ARC_BGC_L4_MY_009_124-TDS --product-id cmems_obs-oc_arc_bgc-plankton_my_l4-multi-1km_P1M --longitude-min 2 --longitude-max 39.2 --latitude-min 72 --latitude-max 85 --date-min "2021-03-01 00:00:00" --date-max "2021-09-01 00:00:00"  --variable CHL   --out-dir <OUTPUT_DIRECTORY> --out-name <OUTPUT_FILENAME> --user <USERNAME> --pwd <PASSWORD> '
```





old script:
python -m motuclient --motu https://my.cmems-du.eu/motu-web/Motu --service-id OCEANCOLOUR_ARC_CHL_L4_REP_OBSERVATIONS_009_088-TDS --product-id dataset-oc-arc-chl-multi_cci-l4-chl_1km_monthly-rep-v02 --longitude-min 2 --longitude-max 39.2 --latitude-min 72 --latitude-max 85 --date-min "2021-03-01 00:00:00" --date-max "2021-09-01 00:00:00"  --variable CHL   --out-dir <OUTPUT_DIRECTORY> --out-name <OUTPUT_FILENAME> --user <USERNAME> --pwd <PASSWORD> 

adjust the script with credentials and other relevant infos

```{r}
cfg <- parse.CMEMS.script(scriptGEN)
cfg2 <- update(cfg,user= user, pwd =pss, out.dir = "data/")
#cfg2
```


# run the download script

define download periods

```{r}
downgrid <- expand.grid(year = c(1998:2021), mmin = 3, mmax = 9)
#downgrid <- rbind(downgrid, data.frame(year = 2022, mmin = 3, mmax = 6))
# downgrid <- rbind(downgrid, data.frame(year = 2021, mmin = 7, mmax = 9))

```

actually run the script and convert it to geotif


```{r}

 for(i in 1:nrow(downgrid)){
    y <- downgrid$year[i]
    mmin <- downgrid$mmin[i]  
    mmax <- downgrid$mmax[i]  
    
    
      
      date.min <- paste0(y, "-0", mmin, "-01 00:00:00")
      date.max <- paste0(y, "-0", mmax, "-01 00:00:00")
      out.name <- paste0("OCEANCOLOUR_ARC_CHL_L4_", y, ".nc")
      longMax <- "50"
      longMin <- "-10"
      latMax <- "85"
      latMin <- "72"
      out.dir <- "data/chl_1km_nc"
      
      
      
      cfg3 <- update(cfg2, 
                     date.min = date.min,
                     date.max = date.max,
                     out.name = out.name,
                     longitude.max = longMax,
                     longitude.min = longMin,
                     latitude.max = latMax,
                     latitude.min = latMin,
                     out.dir = out.dir)
        
      
      
      CMEMS.download(cfg3)
      
      
      br <- brick(here::here(out.dir, out.name))
      setZ(br, as.Date(paste0(y, "-", mmin:mmax, "-01")), "Date")
      
      
      
      writeRaster(br, here::here('data','chl_1km_tif', 
                                 paste0("OCEANCOLOUR_ARC_CHL_L4_", y, ".tif")
      ), overwrite = T)
      
      
      rm(br)
      print(paste(y))
      }

```


andrà integrato con i dati giornalieri

```{r}
downgrid <- expand.grid(year = c(2022), mmin = 3, mmax = 6)


```

actually run the script for 2022 (partial coverage)


```{r}

 for(i in 1:nrow(downgrid)){
    y <- downgrid$year[i]
    mmin <- downgrid$mmin[i]  
    mmax <- downgrid$mmax[i]  
    
    
      
      date.min <- paste0(y, "-0", mmin, "-01 00:00:00")
      date.max <- paste0(y, "-0", mmax, "-01 00:00:00")
      out.name <- paste0("OCEANCOLOUR_ARC_CHL_L4_partial_", y, ".nc")
      longMax <- "50"
      longMin <- "-10"
      latMax <- "85"
      latMin <- "72"
      out.dir <- "data/chl_1km_nc"
      
      
      
      cfg3 <- update(cfg2, 
                     date.min = date.min,
                     date.max = date.max,
                     out.name = out.name,
                     longitude.max = longMax,
                     longitude.min = longMin,
                     latitude.max = latMax,
                     latitude.min = latMin,
                     out.dir = out.dir)
        
      
      
      CMEMS.download(cfg3)
      
      
      br <- brick(here::here(out.dir, out.name))
      setZ(br, as.Date(paste0(y, "-", mmin:mmax, "-01")), "Date")
      
      
      
      writeRaster(br, here::here('data','chl_1km_tif', 
                                 paste0("OCEANCOLOUR_ARC_CHL_L4_partial_", y, ".tif")
      ), overwrite = T)
      
      
      rm(br)
      print(paste(y))
      }

```

