---
title: "0.2_monthly_chla_download"
author: "mz"
format: html
editor: source
---

## Libraries

load the packages needed for downloading the data

```{r}


library(RCMEMS)
library(raster)
```

load credentials


```{r}
# read cred file ingoring row names


cred <- read.csv(here::here('data/credentials.csv'), sep = ";")

user <- cred$user
pss <- cred$pss

```


(raw) Script to download files

cmems_obs-oc_arc_bgc-plankton_nrt_l3-multi-1km_P1D 

```{r}
scriptGEN <- 'python -m motuclient --motu https://nrt.cmems-du.eu/motu-web/Motu --service-id OCEANCOLOUR_GLO_BGC_L4_NRT_009_102-TDS --product-id cmems_obs-oc_glo_bgc-plankton_nrt_l4-gapfree-multi-4km_P1D --longitude-min 7.834812723214284 --longitude-max 28.727646651785708 --latitude-min 70.67026187918526 --latitude-max 84.33096098632811 --date-min "2023-03-07 00:00:00" --date-max "2023-03-07 23:59:59" --variable CHL --variable flags --variable CHL_uncertainty --out-dir <OUTPUT_DIRECTORY> --out-name <OUTPUT_FILENAME> --user <USERNAME> --pwd <PASSWORD> '
```

adjust the script with credentials and other relevant infos

```{r}
cfg <- parse.CMEMS.script(scriptGEN)
cfg2 <- update(cfg,user= user, pwd =pss, out.dir = "data/")
#cfg2
```

## test product

Start the test with two days


```{r}




      date.min <- paste0(2022, "-0", 7, "-01 00:00:00")
      date.max <- paste0(2022, "-0", 7, "-02 00:00:00")
      out.name <- paste0("test_download_", 202205, ".nc")
      longMax <- "50"
      longMin <- "-10"
      #latMax <- "85"
      #latMin <- "72"
      
      latMax <- "60"
      latMin <- "50"

      
      out.dir <- "data/monthly_from_daily"
      
      
      
      cfg3 <- update(cfg2, 
                     date.min = date.min,
                     date.max = date.max,
                     out.name = out.name,
                     longitude.max = longMax,
                     longitude.min = longMin,
                     latitude.max = latMax,
                     latitude.min = latMin,
                     out.dir = out.dir)
        
      
      
      CMEMS.download(cfg3)
      
 

```


the dataset start from sepotember 2022

ems-du.eu/motu-web/Motu
2023-03-09 11:36:01.711 [ERROR] 010-6 : The date range is invalid. Invalid date range: [2022-07-01 00:00:00,2022-07-02 23:59:59]. Valid range is: [2023-03-02 00:00:00,2023-03-07 00:00:00].
[1] 0

*incomprensibile*

in realtà qualcosa dovrebbe esserci

https://marine.copernicus.eu/user-corner/user-notification-service/oc-597-oceancolour-family-products-daily-datasets-incorrectly


provo a scaricarlo con motu

OCEANCOLOUR_ARC_BGC_L3_MY_009_123-TDS



```{r}
scriptGEN <- 'python -m motuclient --motu https://nrt.cmems-du.eu/motu-web/Motu --service-id OCEANCOLOUR_ARC_BGC_L3_MY_009_123-TDS --product-id cmems_obs-oc_arc_bgc-plankton_my_l3-multi-1km_P1D --longitude-min 7.834812723214284 --longitude-max 28.727646651785708 --latitude-min 70.67026187918526 --latitude-max 84.33096098632811 --date-min "2023-03-07 00:00:00" --date-max "2023-03-07 23:59:59" --variable CHL --variable flags --variable CHL_uncertainty --out-dir <OUTPUT_DIRECTORY> --out-name <OUTPUT_FILENAME> --user <USERNAME> --pwd <PASSWORD> '


cfg <- parse.CMEMS.script(scriptGEN)
cfg2 <- update(cfg,user= user, pwd =pss, out.dir = "data/")




      date.min <- paste0(2022, "-0", 7, "-01 00:00:00")
      date.max <- paste0(2022, "-0", 7, "-02 00:00:00")
      out.name <- paste0("test_download_", 202205, ".nc")
      longMax <- "50"
      longMin <- "-10"
      #latMax <- "85"
      #latMin <- "72"
      
      latMax <- "60"
      latMin <- "50"

      
      out.dir <- "data/monthly_from_daily"
      
      
      
      cfg3 <- update(cfg2, 
                     date.min = date.min,
                     date.max = date.max,
                     out.name = out.name,
                     longitude.max = longMax,
                     longitude.min = longMin,
                     latitude.max = latMax,
                     latitude.min = latMin,
                     out.dir = out.dir)
        
      
      
      CMEMS.download(cfg3)
      
 

```

il servizio non sembra essere disponibile, provo a scaricarlo tramite opendap



```{r}
library(ncdf4)
library(lubridate)



server <- "@my.cmems-du.eu";                    # Copernicus Marine server
datasetID <- "cmems_obs-oc_arc_bgc-plankton_my_l3-multi-1km_P1M" # datasetID

USERNAME <- cred$user# readline("Enter your username: ")   # input your username
PASSWORD <- cred$pss# readline("Enter your password: ")   # input your password

```


provo ad accedere al server OPENDAP

```{r}
# creates the OPeNDAP url
url <- paste ("https://",USERNAME, ":", PASSWORD,  server,"/thredds/dodsC/",datasetID, sep = "")

# Open the connection
ds <- nc_open(url)   

ds
```


